\documentclass[a4paper,UTF8]{article}
\usepackage[margin=1.25in]{geometry}
\usepackage{color}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{pdfpages}
\usepackage{fontspec}                   % requires XeLaTeX

\theoremstyle{definition}
\newtheorem*{solution}{Solution}
\newtheorem*{prove}{Proof} 

\setmainfont                    [Ligatures=TeX]{Times New Roman}
\parindent=0pt
\topskip=-200pt

\begin{document}
\title{PTC (Fall 2018) -- Assignment 2}
\author{
    \begin{minipage}[b]{0.3\linewidth}
      \begin{flushright}
        Name: \rule{3cm}{0.4pt}\\
        \vbox to 1mm{}
        Grade: \rule{3cm}{0.4pt}
      \end{flushright}
    \end{minipage}
    \hfill
    \begin{minipage}[b]{0.3\linewidth}
      \begin{flushright}
        Dept: \rule{3cm}{0.4pt}\\
        \vbox to 1mm{}
        ID: \rule{3cm}{0.4pt}
      \end{flushright}
    \end{minipage}
}
\date{\color{red} \Large Due: Nov. 22, 2018}
\maketitle

\section*{Problem 1}
Give context free grammars that generate the following languages, and give a brief description of the functionality of each variable in your grammars (in natural language).
\begin{enumerate}
  \item[a.] \{$w\in \{a,b\}^*\ |\mbox{ the number of $a$'s in $w$ is more than the number of $b$'s in $w$}$\}
  \item[b.] \{$a^ib^{2j}c^{j}d^{k}\ |\ i,j,k\geq 1, k \geq 2i$\}
  \item[c.] \{$w\in \{a,b\}^*\ |\ abb\mbox{ and }aab\mbox{ are substrings of }w$\}
  \item[d.] \{$a^{i}b^{j}c^{k}\ |\ i,j,k\geq 0 \wedge i + j > k$\} 
\end{enumerate}
\begin{solution}
\end{solution}
\newpage

\section*{Problem 2}
Consider the following CFG $G$:
\[
\begin{aligned}
  S &\rightarrow SS\ |\ T \\
  T &\rightarrow aTb\ |\ ab
\end{aligned}
\]
Describe $L(G)$ and show that $G$ is ambiguous. Give an unambiguous grammar $H$
where $L(H) = L(G)$ and sketch a proof that $H$ is unambiguous.
\begin{solution}
\end{solution}
\newpage

\section*{Problem 3}
Let $G = (V, T, P, S)$ be a context free grammar such that $V = \{A, B\}$, $T = \{a, b\}$, $S = B$ and $P$ is
\[
\begin{aligned}
  B &\rightarrow aAA \\
  A &\rightarrow aB \ |\ bB\ |\ a
\end{aligned}
\]
\begin{enumerate}
  \item[a.] Give parse trees, leftmost and rightmost derivations for the following strings.
  \begin{itemize}
    \item[1.] $aabaaa$
    \item[2.] $abaaabaaa$
  \end{itemize}
  \item[b.] Convert $G$ to a PDA that accepts the same language by empty stack.
\end{enumerate}
\begin{solution}
\end{solution}
\newpage

\section*{Problem 4}
Begin with the grammar:
\[
  \begin{aligned}
    S &\rightarrow aAa\ |\ bBb\ |\ \epsilon \\
    A &\rightarrow C \ |\ a \\
    B &\rightarrow C \ |\ b \\
    C &\rightarrow CDE\ |\ \epsilon\\
    D &\rightarrow A\ |\ B\ |\ ab
  \end{aligned}
\]
\begin{enumerate}
  \item[a.] Eliminate $\epsilon$-productions.
  \item[b.] Eliminate any unit productions in the resulting grammar of (a.).
  \item[c.] Eliminate any useless symbols in the resulting grammar of (b.).
  \item[d.] Put the resulting grammar of (c.) into Chomsky normal form.
\end{enumerate}
\begin{solution}
\end{solution}
\newpage

\section*{Problem 5}
Given grammar $G$:
\[
\begin{aligned}
  S & \rightarrow AB\ |\ BC \\
  A & \rightarrow BA\ |\ a \\
  B & \rightarrow CC\ |\ b \\
  C & \rightarrow AB\ |\ a
\end{aligned}
\]
Please use CYK algorithm to decide whether string $ababa$ belongs to $L(G)$.
\begin{prove}
\end{prove}
\newpage

\section*{Problem 6}
Use the CFL pumping lemma to show each of these languages are not context free.
\begin{enumerate}
  \item[a.] $\{a^ib^jc^k\ |\ i < j < k\}$
  \item[b.] $\{0^p\ |\ p\mbox{ is a prime}\}$
  \item[c.] $\{ww^Rw\ |\ w \in \{0,1\}^* \}$
\end{enumerate}
\begin{prove}
\end{prove}
\newpage

\section*{Problem 7}
Show that the CFL is closed under the following operations:
\begin{enumerate}
  \item[a.] $init(L)=\{w\ |\mbox{ for some }x, wx \in L\}$. (\textbf{Hint}: Start with a CNF grammar for the language $L$)
  \item[b.] $cycle(L)=\{xy\ |\ yx \in L\}$ (\textbf{Hint}: Try a PDA-based construction)
\end{enumerate}
\begin{prove}
\end{prove}
\newpage

\section*{Problem 8}
If $L$ is regular, it satisfies pumping lemma for sure! But if a language satisfies pumping lemma, is it regular? Prove or disprove it.
\begin{prove}
\end{prove}
\end{document}
